\hspace{\parindent} 
First and foremost, I would like to thank for all the people who supported me in this long and winding road of life. I would not be standing on the shoulders of the giants without the foundations paved by the people who were behind the scenes of my success.  A special thanks to my advisor Dr. Roland V. Sarmago for never giving up on me, even when I was already on the verge of quitting because of the countless failures and mistakes I've made throughout the journey. I also want to express my gratitude to Dr. Vallerie Samson for the guidance and insights that help me  in my college life.


Thanks also for my friends in Materials Science Society, who have been my first family in UP. To my high school friends, kabarkada, and etc., thank you for all the happy memories. It did an effective job in alleviating my exhaustion in acads and labworks. 

To my lab mentor kuya Jonalds, thank you for guiding me when I was apprentice and still don't know what to do in the lab. Thank you also for the invaluable life tips and all the snacks and foods you brought in. I will return the favor someday. To ate Ches, thank you for all support and reminders even if I was kind stubborn. Thank you also for listening to my problems in life. 

A warm thanks to all my co-lab members,  for the support and advice. A word of thanks to my partner in crime, Lawrence Magsayo. I hope 'magtinarong na ka'. Special mention to Yka, for always there listening to me and reminding me of deadlines and deliverables. 

Finally, I would like to express my gratitude to my mother, father, and sisters for the moral and emotional support (and financial also) that I needed. 

